---
title: "Include hh_size"
author: "Ahmed Javed Malik, Binh Minh An Nguyen"
date: "4/26/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data ETL

```{r package, echo=FALSE}
options(scipen = 4)
library(tidyverse)
library(glmnet)
library(corrplot)
library(boot)
library(broom)
library(car)
library(klaR)
```


```{r bank-acct}
bank_acct <- read_csv("./Train.csv")
```


```{r}
bank_acct_clean <- bank_acct %>% 
  filter(year == 2018) %>% 
  filter(job_type != "Dont Know/Refuse to answer", marital_status != "Dont know") %>% 
  mutate(
    marital_status = case_when(marital_status == "Divorce/Separate" ~ "Divorce",
                               marital_status == "Married/Living together" ~ "Married",
                               marital_status == "Single/Never Married" ~ "Single",
                               TRUE ~ "Widowed"),
    education_level = case_when(
      education_level %in% c("No formal education", "Other/Dont know/RTA") ~ "Others",
      education_level == "Primary education" ~ "Primary School",
      education_level == "Secondary education" ~ "Secondary School",
      TRUE ~ education_level),
    age = age_of_respondent,
    gender = gender_of_respondent,
    bank_account = case_when(bank_account == "Yes" ~ 1,
                             TRUE ~ 0)
  ) %>% 
  dplyr::select(-c(age_of_respondent, gender_of_respondent, year, uniqueid))

# Sampling the Uganda - Government Dependent job_type
RNGkind(sample.kind="default")
set.seed(1)
uganda_gvt <- bank_acct_clean %>% 
  filter(country == "Uganda", job_type == "Government Dependent")
sample <- sample(nrow(uganda_gvt), 0.5 * nrow(uganda_gvt))
uganda_gvt_sample <- uganda_gvt[sample,]

# Add the sample data to bank_account data
bank_acct_smp <- bank_acct_clean %>% rbind(uganda_gvt_sample)

bank_acct_smp <- bank_acct_smp %>% 
  mutate_if(is.character, as_factor)
```


```{r}
bank_acct_smp$relationship_with_head <- relevel(bank_acct_smp$relationship_with_head, "Other non-relatives")
bank_acct_smp$education_level <- relevel(bank_acct_smp$education_level, "Others")
bank_acct_smp$marital_status <- relevel(bank_acct_smp$marital_status, "Single")
bank_acct_smp$country <- relevel(bank_acct_smp$country, "Uganda")
```

# Logistic Regression - with household_size

```{r}
bank.fit <- glm(bank_account~., 
                 data = bank_acct_smp,
                 family = binomial(link = "logit"))
summary(bank.fit)

log.odds <- coef(bank.fit)
odds <- exp(log.odds)
prob <- odds/(1+odds)
cbind(log.odds, odds, prob)

dev.dif <- abs(bank.fit$null.deviance - bank.fit$deviance)
dev.Rsq <- (bank.fit$null.deviance - bank.fit$deviance)/bank.fit$null.deviance
dev.Rsq

cond.index(bank.fit, data = bank_acct_smp)
vif(bank.fit)
```



1. Fit the model on train set

```{r}
RNGkind(sample.kind = "default")
set.seed(1)


tr.size <- 0.7
train <- sample(nrow(bank_acct_smp), tr.size*nrow(bank_acct_smp))
bank.train <- bank_acct_smp[train, ]
bank.test <- bank_acct_smp[-train, ]

fit.train <- glm(bank_account~.,
                 data = bank.train,
                 family = binomial(link = "logit"))
summary(fit.train)

# Compute R-sq - deviance
dev.dif <- abs(fit.train$null.deviance - fit.train$deviance)
dev.Rsq <- (fit.train$null.deviance - fit.train$deviance)/fit.train$null.deviance
dev.Rsq

# Coefficients
log.odds.train <- coef(fit.train)
odds.train <- exp(log.odds.train)
prob.train <- odds.train/(1+odds.train)
cbind(log.odds.train, odds.train, prob.train)

# Train MSE and RMSE
train.mse <- mean(fit.train$residuals ^ 2)
train.rmse <- sqrt(train.mse)
cbind("MSE Train" = train.mse, "RMSE Train" = train.rmse)
```


2. Model validating on test set

```{r}
pred.test <- predict(fit.train, bank.test, type = "response")
head(pred.test)

test.mse.rscv <- mean((bank.test$bank_account - pred.test)^2)
test.rmse.rscv <- sqrt(test.mse.rscv)
cbind("RSCV Test MSE" = test.mse.rscv,
      "RSCV Test RMSE" = test.rmse.rscv)
```

3. Add ROC Curve and Confusion Matrix

```{r}
# Confusion Matrix
thresh <- 0.5
bank.prob.test <-ifelse(pred.test > thresh, 1, 0)
conf.mat <- table("Predicted" = bank.prob.test, "Actual" = bank.test$bank_account)
colnames(conf.mat)<-c("No","Yes")
rownames(conf.mat)<-c("No","Yes")
conf.mat

# ROC Curve
library(ROCR)

pred <- prediction(bank.prob.test, bank.test$bank_account)
perf <- performance(pred, "tpr", "fpr")
plot(perf, colorize = T)

auc <- performance(pred, "auc")
auc.name <- auc@y.name[[1]]
auc.val <- round(auc@y.values[[1]], digits = 3)
paste(auc.name, auc.val)
```

4. CV RMSE - full model

```{r}
test.mse.10f <- cv.glm(bank_acct_smp, bank.fit, K = 10)$delta[1]
test.rmse.10f <- sqrt(test.mse.10f)
cbind("10FCV Test MSE" = test.mse.10f,
      "10FCV Test RMSE" = test.rmse.10f)
```



# Random Forest 

```{r}
library(caret)

as.factor(ifelse(bank_acct_smp$bank_account == 0, "No", "Yes")) -> bank_acct_smp$bank_account
bank_acct_smp$bank_account <- relevel(bank_acct_smp$bank_account, "Yes")

set.seed(1)
bank_acct_smp$bank_account <- as.factor(bank_acct_smp$bank_account)
train <- sample(1:nrow(bank_acct_smp), 0.7*nrow(bank_acct_smp))
bank.rf.train <- bank_acct_smp[train, ]
bank.rf.test<-bank_acct_smp[-train, ]


fitControl1 <- trainControl(method = "repeatedcv",
                            number = 10,
                            search = "random",
                            repeats = 1,
                            savePredictions = T)
modelFitrf <- train(bank_account~.,
                    data = bank.rf.train,
                    method = "rf",
                    trControl = fitControl1,
                    tuneLength = 10,
                    ntree = 100)

modelFitrf$bestTune
plot(varImp(modelFitrf, scale = F), main = "Var Imp: RF 10-FCV")
```

```{r}
sub_rf1=subset(modelFitrf$pred, modelFitrf$pred$mtry==modelFitrf$bestTune$mtry)
caret::confusionMatrix(table(sub_rf1$pred, sub_rf1$obs))
```


**Confusion Matrix for the test set**

```{r}
thresh <- 0.5
pred.prob.rf.50 <- predict(modelFitrf, bank.rf.test, type = "prob")
rf.pred.50 <- ifelse(pred.prob.rf.50[,1] > thresh, 1, 0)


confmat <- table("Predicted" = rf.pred.50, 
                 "Actual" = bank.rf.test$bank_account)
confmat

TruN <- confmat[1,2] # True negatives
TruP <- confmat[2,1] # True positives
FalN <- confmat[1,1] # False negatives
FalP <- confmat[2,2] # False positives
TotN <- TruN + FalP # Total negatives
TotP <- TruP + FalN # Total positives
Tot <- TotN + TotP # Total

c("TruN"=TruN, "TruP"=TruP, "FalN"=FalN, "FalP"=FalP, 
  "TotN"=TotN, "TotP"=TotP, "Tot"=Tot)

Accuracy.Rate <- (TruN + TruP) / Tot
Error.Rate <- (FalN + FalP) / Tot
Sensitivity <- TruP / TotP
Specificity <- TruN / TotN
FalP.Rate <- 1 - Specificity

rf.rates.50 <- c(Accuracy.Rate, Error.Rate, Sensitivity, Specificity, FalP.Rate)

names(rf.rates.50)=c("Accuracy Rate", "Error Rate", 
                       "Sensitivity", "Specificity", "False Positives")
rf.rates.50
```


2. Random forest with Threshold = 0.3

```{r}
thresh <- 0.3
pred.prob.rf.30 <- predict(modelFitrf, bank.rf.test, type = "prob")
rf.pred.30 <- ifelse(pred.prob.rf.30[,1] > thresh, 1, 0)

confmat <- table("Predicted" = rf.pred.30, "Actual" = bank.rf.test$bank_account) 
confmat

TruN <- confmat[1,2] # True negatives
TruP <- confmat[2,1] # True positives
FalN <- confmat[1,1] # False negatives
FalP <- confmat[2,2] # False positives
TotN <- TruN + FalP # Total negatives
TotP <- TruP + FalN # Total positives
Tot <- TotN + TotP # Total

c("TruN"=TruN, "TruP"=TruP, "FalN"=FalN, "FalP"=FalP, 
  "TotN"=TotN, "TotP"=TotP, "Tot"=Tot)

Accuracy.Rate <- (TruN + TruP) / Tot
Error.Rate <- (FalN + FalP) / Tot
Sensitivity <- TruP / TotP
Specificity <- TruN / TotN
FalP.Rate <- 1 - Specificity

rf.rates.30 <- c(Accuracy.Rate, Error.Rate, Sensitivity, Specificity, FalP.Rate)

names(rf.rates.30)=c("Accuracy Rate", "Error Rate", 
                       "Sensitivity", "Specificity", "False Positives")
rf.rates.30
```


2. Random forest with Threshold = 0.7

```{r}
thresh <- 0.7
pred.prob.rf.70 <- predict(modelFitrf, bank.rf.test, type = "prob")
rf.pred.70 <- ifelse(pred.prob.rf.70[,1] > thresh, 1, 0)

confmat <- table("Predicted" = rf.pred.70, "Actual" = bank.rf.test$bank_account) 
confmat

TruN <- confmat[1,2] # True negatives
TruP <- confmat[2,1] # True positives
FalN <- confmat[1,1] # False negatives
FalP <- confmat[2,2] # False positives
TotN <- TruN + FalP # Total negatives
TotP <- TruP + FalN # Total positives
Tot <- TotN + TotP # Total

c("TruN"=TruN, "TruP"=TruP, "FalN"=FalN, "FalP"=FalP, 
  "TotN"=TotN, "TotP"=TotP, "Tot"=Tot)

Accuracy.Rate <- (TruN + TruP) / Tot
Error.Rate <- (FalN + FalP) / Tot
Sensitivity <- TruP / TotP
Specificity <- TruN / TotN
FalP.Rate <- 1 - Specificity

rf.rates.70 <- c(Accuracy.Rate, Error.Rate, Sensitivity, Specificity, FalP.Rate)

names(rf.rates.70)=c("Accuracy Rate", "Error Rate", 
                       "Sensitivity", "Specificity", "False Positives")
rf.rates.70
```

**Combine**

```{r}
AllStats <- round(rbind(rf.rates.30,
                        rf.rates.50,
                        rf.rates.70),
                  digits = 3)
rownames(AllStats) <- c("Threshold = 30%",
                        "Threshold = 50%",
                        "Threshold = 70%")
AllStats
```


### ROCR Curve for Random Forest (on the test set)

```{r}
library(ROCR)

pred.rf <- prediction(rf.pred.50, bank.rf.test$bank_account) 
perf.rf <- performance(pred.rf, "tpr", "fpr")

plot(perf.rf, colorize = T)

auc.rf <- performance(pred.rf,"auc") # Compute the AUC

auc.name.rf <- auc.rf@y.name[[1]] # Retrieve the AUC label text
auc.value.rf <- round(auc.rf@y.values[[1]], digits = 3) # Retrieve the AUC value, rounded

paste(auc.name.rf, "is", auc.value.rf)
```


### CV for Random Forest

```{r}
library(caret)

set.seed(1)

fitControl1 <- trainControl(method = "repeatedcv",
                            number = 10,
                            search = "random",
                            repeats = 1,
                            savePredictions = T)
modelFitrf <- train(bank_account~.,
                    data = bank_acct_smp,
                    method = "rf",
                    trControl = fitControl1,
                    tuneLength = 10,
                    ntree = 100)

modelFitrf$bestTune
plot(varImp(modelFitrf, scale = F), main = "Var Imp: RF 10-FCV")
```

```{r}
sub_rf1=subset(modelFitrf$pred, modelFitrf$pred$mtry==modelFitrf$bestTune$mtry)
caret::confusionMatrix(table(sub_rf1$pred, sub_rf1$obs))
```

# Boosted Tree

```{r}
library("gbm")

set.seed(1)

tr.size <- 0.7
train <- sample(1:nrow(bank_acct_smp), 
                tr.size * nrow(bank_acct_smp)) # Train index

bank.boost.train <- bank_acct_smp[train,]
bank.boost.test <- bank_acct_smp[-train,] 

bank.boost.train$bank_account <- as.numeric(bank.boost.train$bank_account)
boost.bank <- gbm(bank_account~.,
                  data = bank.boost.train,
                  distribution = "bernoulli",
                  shrinkage = 0.01,
                  cv.folds = 10,
                  n.trees = 3500, 
                  interaction.depth = 1)
boost.bank

# Number of boosted trees with the smallest CV test error
best.num.trees <- which.min(boost.bank$cv.error)
best.num.trees

# The smallest CV Test Error
min.10FCV.error <- round(min(boost.bank$cv.error), digit=4)
min.10FCV.error

# Display result
paste("Min 10-FCV Test Error" = min.10FCV.error, "at" , best.num.trees,"trees")
summary(boost.bank)

# Plotting the graph
plot(boost.bank$cv.error, type = "l", xlab = "Number of Trees", ylab = "CV Test MSE")
```

### Confusion matrix for Boosted Tree

```{r}
# Threshold = 0.5
thresh <- 0.5
pred.boost.50 <- predict(boost.bank, bank.boost.test, type = "response")
boost.pred.prob.50 <- ifelse(pred.boost.50 > thresh, 1, 0)

confmat <- table("Predicted" = boost.pred.prob.50, "Actual" = bank.boost.test$bank_account) 
confmat

TruN <- confmat[1,1] # True negatives
TruP <- confmat[2,2] # True positives
FalN <- confmat[1,2] # False negatives
FalP <- confmat[2,1] # False positives
TotN <- TruN + FalP # Total negatives
TotP <- TruP + FalN # Total positives
Tot <- TotN + TotP # Total

c("TruN"=TruN, "TruP"=TruP, "FalN"=FalN, "FalP"=FalP, 
  "TotN"=TotN, "TotP"=TotP, "Tot"=Tot)

Accuracy.Rate <- (TruN + TruP) / Tot
Error.Rate <- (FalN + FalP) / Tot
Sensitivity <- TruP / TotP
Specificity <- TruN / TotN
FalP.Rate <- 1 - Specificity

boost.rates.50 <- c(Accuracy.Rate, Error.Rate, Sensitivity, Specificity, FalP.Rate)

names(boost.rates.50)=c("Accuracy Rate", "Error Rate", 
                       "Sensitivity", "Specificity", "False Positives")
boost.rates.50
```


**Threshold = 0.3**

```{r}
thresh <- 0.3
pred.boost.30 <- predict(boost.bank, bank.boost.test, type = "response")
boost.pred.prob.30 <- ifelse(pred.boost.30 > thresh, 1, 0)

confmat <- table("Predicted" = boost.pred.prob.30, "Actual" = bank.boost.test$bank_account) 
confmat

TruN <- confmat[1,1] # True negatives
TruP <- confmat[2,2] # True positives
FalN <- confmat[1,2] # False negatives
FalP <- confmat[2,1] # False positives
TotN <- TruN + FalP # Total negatives
TotP <- TruP + FalN # Total positives
Tot <- TotN + TotP # Total

c("TruN"=TruN, "TruP"=TruP, "FalN"=FalN, "FalP"=FalP, 
  "TotN"=TotN, "TotP"=TotP, "Tot"=Tot)

Accuracy.Rate <- (TruN + TruP) / Tot
Error.Rate <- (FalN + FalP) / Tot
Sensitivity <- TruP / TotP
Specificity <- TruN / TotN
FalP.Rate <- 1 - Specificity

boost.rates.30 <- c(Accuracy.Rate, Error.Rate, Sensitivity, Specificity, FalP.Rate)

names(boost.rates.30)=c("Accuracy Rate", "Error Rate", 
                       "Sensitivity", "Specificity", "False Positives")
boost.rates.30
```


**Threshold = 0.7**

```{r}
thresh <- 0.7
pred.boost.70 <- predict(boost.bank, bank.boost.test, type = "response")
boost.pred.prob.70 <- ifelse(pred.boost.70 > thresh, 1, 0)

confmat <- table("Predicted" = boost.pred.prob.70, "Actual" = bank.boost.test$bank_account) 
confmat

TruN <- confmat[1,1] # True negatives
TruP <- confmat[2,2] # True positives
FalN <- confmat[1,2] # False negatives
FalP <- confmat[2,1] # False positives
TotN <- TruN + FalP # Total negatives
TotP <- TruP + FalN # Total positives
Tot <- TotN + TotP # Total

c("TruN"=TruN, "TruP"=TruP, "FalN"=FalN, "FalP"=FalP, 
  "TotN"=TotN, "TotP"=TotP, "Tot"=Tot)

Accuracy.Rate <- (TruN + TruP) / Tot
Error.Rate <- (FalN + FalP) / Tot
Sensitivity <- TruP / TotP
Specificity <- TruN / TotN
FalP.Rate <- 1 - Specificity

boost.rates.70 <- c(Accuracy.Rate, Error.Rate, Sensitivity, Specificity, FalP.Rate)

names(boost.rates.70)=c("Accuracy Rate", "Error Rate", 
                       "Sensitivity", "Specificity", "False Positives")
boost.rates.70
```


**Combine**

```{r}
AllStats <- round(rbind(boost.rates.30,
                        boost.rates.50,
                        boost.rates.70),
                  digits = 3)
rownames(AllStats) <- c("Threshold = 30%",
                        "Threshold = 50%",
                        "Threshold = 70%")
AllStats
```



### ROCR Curve Boosted Tree

```{r}
# ROCR Curve
library(ROCR)

pred.boost <- prediction(boost.pred.prob.50, bank.boost.test$bank_account) 
perf.boost <- performance(pred.boost, "tpr", "fpr")

plot(perf.boost, colorize = T)

auc.boost <- performance(pred.boost,"auc") # Compute the AUC

auc.name.boost <- auc.boost@y.name[[1]] # Retrieve the AUC label text
auc.value.boost <- round(auc.boost@y.values[[1]], digits = 3) # Retrieve the AUC value, rounded

paste(auc.name.boost, "is", auc.value.boost)
```





## CV for Boosted Tree (whole dataset)

```{r}
bank_acct_smp$bank_account <- as.numeric(bank_acct_smp$bank_account)
boost.bank <- gbm(bank_account~.,
                  data = bank_acct_smp,
                  distribution = "bernoulli",
                  shrinkage = 0.01,
                  cv.folds = 10,
                  n.trees = 3500, 
                  interaction.depth = 1)
boost.bank

# Confusion Matrix
thresh <- 0.5
  
pred.boost.full <- predict(boost.bank, bank_acct_smp, type = "response")
boost.pred.prob.full <- ifelse(pred.boost.full > thresh, 1, 0)

confmat <- table("Predicted" = boost.pred.prob.full, "Actual" = bank_acct_smp$bank_account) 
confmat

TruN <- confmat[1,1] # True negatives
TruP <- confmat[2,2] # True positives
FalN <- confmat[1,2] # False negatives
FalP <- confmat[2,1] # False positives
TotN <- TruN + FalP # Total negatives
TotP <- TruP + FalN # Total positives
Tot <- TotN + TotP # Total

c("TruN"=TruN, "TruP"=TruP, "FalN"=FalN, "FalP"=FalP, 
  "TotN"=TotN, "TotP"=TotP, "Tot"=Tot)

Accuracy.Rate <- (TruN + TruP) / Tot
Error.Rate <- (FalN + FalP) / Tot
Sensitivity <- TruP / TotP
Specificity <- TruN / TotN
FalP.Rate <- 1 - Specificity

boost.rates.full <- c(Accuracy.Rate, Error.Rate, Sensitivity, Specificity, FalP.Rate)

names(boost.rates.full)=c("Accuracy Rate", "Error Rate", 
                       "Sensitivity", "Specificity", "False Positives")
boost.rates.full
```



# APPENDIX

## Stepwise for logistic regression

```{r}
bank.null <- lm(bank_account~1, data = bank_acct_smp, family = binomial(link = "logit"))

bank.fit.step <- step(bank.fit, 
                      scope = list(lower = bank.null, 
                                   upper = bank.fit), 
                      direction = "both", 
                      test = "F")

summary(bank.fit.step)
```

